---
title: "Sleep Case Study"
author: "Sarah Schwartz"
date: "12/22/2021"
output: html_document
---



## Sleep Study Report

This is an example of a sleep report template using R Markdown. This template 
will automatically update all plots and statistics if new data if added. 


```{r installing libraries, inlcude=FALSE, echo=FALSE, results='hide', message=FALSE}
packages <- c("tidyverse", "reshape2", "ggpubr")
install.packages(setdiff(packages, rownames(installed.packages())),
                 repos = "http://cran.us.r-project.org")
```


```{r loading libraries, inlcude = FALSE, echo=FALSE, results='hide', message=FALSE}
library(tidyverse)
library(reshape2)
library(ggpubr)
```


```{r pulling in data, inlcude = FALSE, echo=FALSE, message=FALSE}
sleep_data <- read_csv(file = "../Data/sleep_data.csv")
```

``` {r adding columns, inlcude = FALSE, message=FALSE, echo=FALSE}
sleep_data <- sleep_data %>%
  add_column(sleep_month = months(sleep_data$sleep_date))
sleep_data <- sleep_data %>%
  add_column(sleep_quarter = quarters(sleep_data$sleep_date))
```



```{r long form data, inlcude=FALSE, message=FALSE, echo=FALSE}
sleep_data_long <- melt(sleep_data, id.vars = c("sleep_date", "sleep_month",
                                                "sleep_quarter", "duration_hrs",
                                                "total_toss_turns"))
```

## Sleep Duration Stastistics
The following table report the mean, standard deviation and variance for the duration of sleep for the individual. This includes 
``` {r duration Statistics, message=FALSE, echo=FALSE}
ave_duration <- mean(sleep_data$duration_hrs)
std_duration <- sqrt(var(sleep_data$duration_hrs))
var_duration <- var(sleep_data$duration_hrs)
#stat data frame
stats <- data.frame(Statistics = c("Mean", "Standard Deviation", "Variance"),
                    Results = c(round(ave_duration, digits = 2),
                                round(std_duration, digits = 2),
                                round(var_duration, digits = 2)))

knitr::kable(stats)
```


# Plots

## Plotting a bar plot of sleep duration with different colors for the months
```{r bar plot - sleep duration, echo=FALSE}
sleep_data %>%
  ggplot() +
  geom_bar(data = sleep_data,
           mapping = aes(x = duration_hrs, color = sleep_month)) +
  xlab("Sleep Duration (hrs)") +
  ylab("Frequency") +
  ggtitle("Sleep Duration")
```


## Histogram with Normal Distribution Overlay
This is an example of fitting a normal distribution with mean of and standard deviation of to the histogram of the sleep duration. 
```{r plot: histtogram with normal distribution, echo=FALSE}
dur <- sleep_data$duration_hrs
h <- hist((sleep_data$duration_hrs), main = "Sleep Duration",
          xlab = "Sleep Duration (hrs)")

xfit <- seq(min(dur), max(dur), length = 40)
yfit <- dnorm(xfit, mean = mean(dur), sd = sd(dur))
yfit <- yfit * diff(h$mids[1:2]) * length(dur)

lines(xfit, yfit, col = "blue", lwd = 2)
```


## Testing Normality of the data using the Shapiro-Wilk Test
This test will determine of the duration of sleep is consistent with a normal distribution.
The hypothesis is that the data is consistent with a normal distribution.
The results from the calculation is a p-value of 4.199e-10. Although the data may appear to be normally distributed when plotted as a histogram, the statistical test indicate that it is not consistent. This variation from the normal distribution, however, is much clearer when plotted in a Q-Q plot, as seen below.

```{r plot: qqplot, echo=FALSE}
ggqqplot(sleep_data$duration_hrs)
```
